---
title: "Time Spent in Meetings,  Mining Outlook Calendar"
author: "Ann Crawford"
date: "April 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## A Simple,Free Tool for Time Management

**How much time to you spend in meetings?** Transparent data analysis of outlook calendar inspired by Tableau video <http://inn0vate.blogspot.com/2012/09/setting-up-csv-file-for-outlook.html>.

Open outlook, Click file>  Open and Export. Select Export to a file > next > Select Comma separated Values > next > Select Calendar > save file.

##Data Processing
```{r readandprocessdata}
##install.packages("ggplot2")
##install.packages("dplyr")
##install.packages("lubridate")

rm(list = ls())
library(ggplot2)
library(dplyr)
library(lubridate)

filefullpath <- "D:/Temp/MeetingTest.csv"
if (!file.exists(filefullpath) )
{
meetingraw <-  read.csv(filefullpath)
}

meeting <- select(meetingraw, Subject, Start.Date, Start.Time, End.Date, End.Time, Location,
        Show.time.as, All.day.event, Private ) 

## use libridate function to create data class variable yyyy-mm-dd hh:mm:ss with 24 hour clock
meeting$StartDateTime <- mdy_hms(paste(meeting$Start.Date, meeting$Start.Time, sep = ":" ) )
meeting$EndDateTime   <- mdy_hms(paste(meeting$End.Date, meeting$End.Time, sep = ":" ) )
                                 

```

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
