---
title: "Time Spent in Meetings, Mining Outlook Calendar"
author: "Ann Crawford"
date: "April 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## A Simple,Free Tool for Time Management

**How much time to you spend in meetings?** Transparent data analysis of outlook calendar inspired by Tableau video <http://inn0vate.blogspot.com/2012/09/setting-up-csv-file-for-outlook.html>.

Open outlook, Click file>  Open and Export. Select Export to a file > next > Select Comma separated Values > next > Select Calendar > save file.

##Data Read and Clean Up
```{r readandprocessdata}
##install.packages("ggplot2")
##install.packages("dplyr")
##install.packages("lubridate")

rm(list = ls())
library(ggplot2)
library(dplyr)
library(lubridate)

showtime <- function(i)
{
        ## showtime as 1 = tenative, 2 = busy, 3 = free, 4 = out of office
        result <-       ifelse(i == 1, "tenative", 
                        ifelse(i == 2, "busy" ,
                        ifelse(i == 3, "free",
                        ifelse(i == 4, "out of office"
                               ,"other") ) ) )
        return(result)
}


filefullpath <- "D:/Temp/MeetingTest.csv"
if (file.exists(filefullpath) )
{
meetingraw <-  read.csv(filefullpath)
}


## use libridate function to create data class variable yyyy-mm-dd hh:mm:ss with 24 hour clock
meetingraw$StartDateTime <- mdy_hms(paste(meetingraw$Start.Date, meetingraw$Start.Time, sep = ":" ) )
meetingraw$EndDateTime   <- mdy_hms(paste(meetingraw$End.Date, meetingraw$End.Time, sep = ":" ) )
 
## Get meetings that were actually accepted 
meeting <-   filter(meetingraw, Show.time.as == 2) %>%
select( Subject, StartDateTime , EndDateTime , Location, Show.time.as , All.day.event,                                     Private)   %>% 
        mutate( nbrofminutes = difftime(EndDateTime, StartDateTime, units = "mins")  , weekday = weekdays.Date(StartDateTime)  , month = months.Date(StartDateTime)  ) 
                       

firstday <- date(min(meeting$StartDateTime))
lastday  <- date(max(meeting$StartDateTime))

```
**From `r firstday` to `r lastday` you have spent `r sum(meeting$nbrofminutes)/60` hours in meetings.**
```{r cars}
##summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
##plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
